---
title: "PA1_template.RMD"
author: "Amarendra"
date: "July 8, 2017"
output: 
  html_document:
    keep_md: true 
---
## Initial set up
```{r}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(Hmisc)
library(ggplot2)
```

## Load and Preprocess the data 
```{r}
activityData <- read.csv("activity.csv", header=TRUE,stringsAsFactors=FALSE)
activityData$date<-ymd(activityData$date)
str(activityData)
```

## What is mean total number of steps taken per day?
1. Histogram
```{r}
steps <- activityData%>%filter(!is.na(steps))%>%group_by(date)%>%
  summarise(steps=sum(steps))
head(steps)
hist(steps$steps,col="yellow",
 xlab="Total number of steps",
 main="Histogram of steps taken per day")
```

2. Mean and median of the total number of steps taken per day
```{r}
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)
mean_steps
median_steps
```
* Mean:`r mean_steps`
* Median:`r median_steps`

## What is the average daily activity pattern?
```{r}
interval <- activityData %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarise(steps = mean(steps))
```

1. Timeseries plot
```{r}
ggplot(data=interval, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps<-interval[which.max(interval$steps),]
```
* Most steps are at interval `r maxSteps`

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset
```{r}
missingVal<-sum(is.na(activityData$steps))
```
* Number of missing values: `r missingVal`

2. Fill in the missing values with average number of steps in the same 5-min interval.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```
4. Make a histogram of the total number of steps taken each day
```{r}
stepsByDayImputed <- activityDataImputed%>%group_by(date)%>%
  summarise(steps=sum(steps))
head(stepsByDayImputed)
hist(stepsByDayImputed$steps,col="blue",
     xlab="Total number of steps per day(Imputed)",
     main="Histogram of steps taken per day(Imputed)")
```

5. Calculate and report the mean and median total number of steps taken per day.
```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed$steps)
stepsByDayMedianImputed <- median(stepsByDayImputed$steps)
```
* Mean(Imputed): `r stepsByDayMeanImputed`
* Median(Imputed): `r stepsByDayMedianImputed`

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
averages <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averages, aes(interval, steps)) + 
   geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("5-minute interval") + 
  ylab("Number of steps")
```


